Predicting solutions for the Military Flight and Maintenance Planning problem

### 2020-09-09 R/presentations/EDAA202003/presentation.Rmd
more changes EDAA.

---
title: "Predicting solutions for the Military Flight and Maintenance Planning problem"
author: |
    | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
    | \tiny (franco.peschiera@isae-supaero.fr)
date: "September, 2020"
output:
  xaringan::moon_reader:
    css: ["footer-header.css", "default"]
    self_contained: false
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
  slidy_presentation:
    css: ./../../../css/baobab_slidy.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    toc: true
  beamer_presentation:
    theme: "Berlin"
    toc: true
    slide_level: 2
    incremental: yes
    fig_height: 4
    includes:
        in_header: header.tex
editor_options: 
  chunk_output_type: console
---

@@

1. Problem.
2. State of the art.
1. Solving techniques.
2. Machine Learning.
2. Results.
3. Conclusions.

@@

### Problem (informally)

Assign both **missions** and **maintenance operations** to a fleet of aircraft in order to maximize availability and minimize costs. **Missions** have fixed start and end times and have particular needs in terms of aircraft and flight hours. **Maintenance operations** have specific rules that govern their frequency and capacity.

.center[
```{r, echo=FALSE, out.width='100%', warning=FALSE}
knitr::include_graphics('calendar3.png')

```
]

### Military aircraft maintenance

@@

* Mirage 2000 fleet.
* A single workshop to do the maintenance operations.
* Check type D (overhaul maintenances): 
  * every 5 years or 1000 flight hours and 
  * takes 4-6 months.
* Fleets of 15-60 aircraft.
* Planning horizons of 90 periods (= 7.5 years).

@@

* A series of $j \in \mathcal{J}$ pre-scheduled missions are planned along a horizon divided into $t \in \mathcal{T}$ periods.
* Each mission requires a certain **number** $R_{j}$ of aircraft $i \in \mathcal{I}$ which it employs for $H_j$ **hours** in each period.
* The heterogeneous fleet requires recurrent preventive maintenance operations (**checks**).
* A check takes exactly $M^d$ periods and cannot be interrupted. There is a limited **capacity** for maintenances at each period.
* The objective is to **minimize the total number of checks** while **maximizing the status of the fleet** at the each period of the horizon.

**Other**: soft constraints on the state of the fleet at each period, initial conditions, default consumption, reduce variance of frequency of checks.

@@

* FMP: Flight and Maintenance Planning problem.
* In `r Citep(myBib, "Cho2011")`, US Army aircraft were assigned daily operations over a year to aircraft in order to minimize the maximum number of maintenances.
* In `r Citep(myBib, "Kozanidis2008")`, Greek aircraft had monthly assignments of maintenances and flight hours in order to maximize the availability and final state of squadrons.
* In `r Citep(myBib, "Verhoeff2015")`, monthly assignments were done and several objectives were taken into account: availability, serviceability and final state.
* In `r Citep(myBib, "Seif2018")`, a generalization for different types of maintenances and capacities was done.
* In `r Citep(myBib, "Peschiera2019")`, the problem was proved NP-Complete.

@@

$a_{ijt_1t_2}$: mission $j$ is assigned to aircraft $i$ between $t_1$ and $t_2$.  
$m_{it_1t_2}$: aircraft $i$ starts a check in $t_1$ and another check in $t_2$.

\begin{align}
  & \text{Max}\;
  \sum_{i \in \mathcal{I}, p \in \mathcal{P}} m_{ip} \times W_p \\
  & \sum_{i \in \mathcal{I}, p \in \mathcal{P}_{t}} m_{ip} \leq C^{max} 
          & t \in \mathcal{T} \label{eq:capacity1}\\
  & \sum_{i \in \mathcal{I}_j, (t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \geq R_j
          & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
  & \sum_{p \in \mathcal{P}_{t}} m_{ip} + \sum_{j \in \mathcal{J}_t \cap \mathcal{J}_i} \sum_{(t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \leq 1 
          & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}\\
  & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^M + M (1 - m_{ip}) & \notag \\
    & & i \in \mathcal{I}, p \in \mathcal{P}, c \in \mathcal{C}_p \\
\end{align}

@@

### Forecasting + Optimization

**We want to**:

1. Train a statistical model to predict characteristics of good (optimal?) solutions for any given instance.
2. Use this information to limit the solution space and make it easier to find those potentially good solutions.

@@

**Benefits**:

1. **Performance**: a smaller problem is easier to solve.
2. **User feedback**: direct feedback about the solution without needing to solve any model.
3. **More stable solutions**: Every aircraft flies an amount that is closest to the mean of the fleet.

**The better we're able to predict those characteristics for the whole fleet, the less optimality we will lose**

@@

### Prediction model

* **Technique**: *Quantile regressions* to estimate upper and lower bounds.
* **Training**: 5000 small instances.
* **Input features**: 
  * mean flight demand per period,
  * total remaining flight hours at start (init), 
  * variance of flight demand, 
  * demand of special missions, 
  * number of period where flight demand is cut in two.
* **Output features**: mean distance between maintenances.

### How good is it (performance)

Faster solutions, more solutions.

.center[
```{r, echo=FALSE, out.width='80%'}
imgs = c('time_performance_ordered_2tasks.png')
knitr::include_graphics(imgs)
```
]

### How good is it (optimality)

For instances were an optimal solution was found (optimum degradation):
* 95% of instances had less than 4% gap with real optimal.

.center[
```{r, echo=FALSE, out.width='80%'}
imgs = c('quality_degradation_2tasks.png')
knitr::include_graphics(imgs)
```
]

@@

* **Better predictions** with better features, or predicting several characteristics of optimal solutions.
* **Predict a distribution** and sample patterns from the distribution instead of predicting patterns.
* **Warm-start Column Generation** with a selected subset of potentially good patterns.
* **Automatize prediction** so it can be easily integrated in other problems.

@@

* **Dasault Aviation**: desktop software OPTIMA built, tested, deployed and validated by users.

*Estimated thesis defense: Novembre 2020.*

>>>>

---
title: "Predicting solutions for the Military Flight and Maintenance Planning problem"
author: |
    | Franco Peschiera
    | \tiny (franco.peschiera@isae-supaero.fr)
date: "September, 2020"
output:
  xaringan::moon_reader:
    css: ["footer-header.css", "default"]
    self_contained: false
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
  slidy_presentation:
    css: ./../../../css/baobab_slidy.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    toc: true
  beamer_presentation:
    theme: "Berlin"
    toc: true
    slide_level: 2
    incremental: yes
    fig_height: 4
    includes:
        in_header: header.tex
editor_options: 
  chunk_output_type: console
---

@@

1. Problem.
2. State of the art.
1. Solving techniques.
2. Machine Learning + Optimization.
2. Results.
3. Future work.
3. Publications.

@@

### Military aircraft operations

@@

* Mirage 2000 fleet.
* Fleets of 60-80 aircraft.
* A single workshop to do the maintenance operations.
* Check type D (overhaul maintenances): 
  * every 5 years or 1000 flight hours and 
  * takes 4-6 months.
* Planning horizons of 90 periods (= 7.5 years).
* Missions require:
  * aircraft.
  * monthly flight hours.

@@

A good solution:

* complies with all regulatory rules and mission demands.
* uses the least possible number of maintenances.
* keeps the fleet in a "good state" at every period.

@@

* FMP: Flight and Maintenance Planning problem.
* In `r Citep(myBib, "Cho2011")`, US Army aircraft were assigned daily operations over a year to aircraft in order to minimize the maximum number of maintenances.
* In `r Citep(myBib, "Kozanidis2008")`, Greek aircraft had monthly assignments of maintenances and flight hours in order to maximize the availability and final state of squadrons.
* In `r Citep(myBib, "Verhoeff2015")`, monthly assignments were done and several objectives were taken into account: availability, serviceability and final state.
* In `r Citep(myBib, "Seif2018")`, a generalization for different types of maintenances and capacities was done.

@@

$a_{ijt_1t_2}$: mission $j$ is assigned to aircraft $i$ between $t_1$ and $t_2$.  
$m_{ip}$: aircraft $i$ has a maintenance pattern $p$.

\begin{align}
  & \text{Min}\;
  \sum_{i \in \mathcal{I}, p \in \mathcal{P}} m_{ip} \times W_p \\
  & \sum_{i \in \mathcal{I}, p \in \mathcal{P}_{t}} m_{ip} \leq C^{max} 
          & t \in \mathcal{T} \label{eq:capacity1}\\
  & \sum_{i \in \mathcal{I}_j, (t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \geq R_j
          & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
\end{align}

@@

### Machine Learning + Optimization

We:

1. Train a statistical model to predict certain characteristics of good (optimal?) solutions for 5000 simulated instances.
2. Use this information to *predict* the characteristics of good solutions on a new instance ($\hat{\mu}_{t'-t}^{lb}$,  $\hat{\mu}_{t'-t}^{ub}$).
3. Apply the prediction to *guide* de solution of said instance.

@@

\begin{align}
    & m_{ip} = 0 & p_{t'} - p_t < \hat{\mu}_{t'-t}^{lb} - tol  \\
    & m_{ip} = 0  &  p_{t'} - p_t > \hat{\mu}_{t'-t}^{ub} + tol
\end{align}

@@

## Results

* 30% reduction in solution times.
* 40% more stable solutions.
* 3-4% optimality loss.

@@

* **Better predictions** with better features, or predicting several characteristics of optimal solutions.
* **Warm-start Column Generation** with a selected subset of potentially good patterns.
* **Automatize prediction** so it can be easily integrated in other problems.

@@

* **Dasault Aviation**: desktop software built, tested, deployed and validated by users.
